---
title: "The Socialist Brand in America"
description: |
  A short description of the post.
author:
  - name: Jan Zilinsky
    url: {}
date: 02-21-2021
output:
  distill::distill_article:
    self_contained: false
    code_folding: true
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(haven)
library(pollster)
library(here)

source(here("files","paths.R"))
source("~/Dropbox/JZ-tools.R")
```

In a recent interview with Noah Smith, [David Shor was asked](https://noahpinion.substack.com/p/video-interview-david-shor-political) whether it was true that Hispanic people voted Republican in greater proportions compared to earlier elections. And the harder question was: why?

Shor mentioned possible explanations: 

* An association with socialism is likely to have hurt the Democratic brand among Hispanics.
* Attitudes on law-and-order issues are fairly conservative among Hispanics. 

Here I'll use 50 Nationscape waves to look at average ratings of "socialists" among Hispanic and non-Hispanic respondents across the ideological spectrum.

```{r Import NS data}
# Download data from https://www.voterstudygroup.org/publication/nationscape-data-set
# [for to replicate the plots below, pool all NS data from waves 1 through 50]
a <- read_dta(nationscape_path) %>%
    mutate(ideo3 = case_when(ideo5 <=2 ~ "liberals",
                           ideo5 == 3 ~ "moderates",
                           ideo5 %in% c(4,5) ~ "conservatives"))
```

There are `r a %>% filter(Hispanic==1) %>% tally()` Hispanic respondents in the dataset, and about a half (`r a %>% filter(Hispanic==1, White==1) %>% tally()`) also identify as White. Among all Hispanics, self-reported ideology is available for N = `r a %>% filter(Hispanic==1, !is.na(ideo3)) %>% tally()` respondents.

```{r}
ratings_S <- a %>% 
  filter(group_favorability_socialists != 999, !is.na(ideo3)) %>%
  group_by(ideo3,Hispanic, White) %>%
  filter(Hispanic == 1 | White ==1) %>%
  summarise(n = n(),
            M = weighted.mean(group_favorability_socialists, na.rm=T)) %>%
  mutate(White = recode(White, `0` = "Non-White", `1` = "White"),
         Hispanic = recode(Hispanic, `0` = "Non-Hispanic", `1` = "Hispanic")) %>%
  unite(var, White, Hispanic, ideo3, sep = " ")

ratings_S %>%
  ggplot(aes(y=reorder(var,M), x=M)) + geom_point(size=2) +
  theme_bw() + theme(text = element_text(size=11.5)) +
  labs(title = "Ratings of socialists among Nationscape respondents") +
  labs(x= "1-4 scale (2 = Somewhat favorable, 3 = Somewhat unfavorable)", y = "")
```

Here averages and samples sizes for the same groups:

```{r}
ratings_S %>%
  rename(Group = var, 
         `Number of respondents` = n,
         `Avg. ratings of socialists` = M) %>%
  arrange(`Avg. ratings of socialists`) %>%
  relocate(Group, `Avg. ratings of socialists`) %>%
knitr::kable(digits = 2)
```

Rather than averaging the 1-4 scale, let's just the display % unfavorable + % very unfavorable:

```{r}
a <- a %>%
  mutate(socialists_UNFAV = case_when(group_favorability_socialists <= 2 ~ 0,
                                group_favorability_socialists %in% c(3,4) ~ 1))

unfavTOplot <- moe_crosstab(df = a %>% filter(hispanic %in%  c(1:5,12:15),
                                              intendsToVote == 1,
                                              registered==1),
         x = hispanic,
         y = socialists_UNFAV,
         weight = weight, n = F, unwt_n = T) %>%
  filter(socialists_UNFAV==1)

unfavTOplot %>%
  ggplot(aes(y=fct_reorder(hispanic,pct), x = pct)) +
  geom_pointrange(aes(xmin = pct-moe, xmax = pct+moe), size=.3) +
  xlim(c(20,80)) +
  labs(x = "Percent", y = "",
       title = "Unfavorable views of socialists among registered likely voters",
       subtitle = "Hispanic respondents broken down by ancestry",
       caption = paste("Data: Nationscape, Waves 1-50. N = ",
                       scales::comma(sum(unfavTOplot$unweighted_n)))) +
  theme_minimal()
```

```{r}
############################################################################
# Prepare a set of age_group-x-pid3 tables for each Hispanic origin group
############################################################################
TP <- NULL
for (hisp in c(1:5,12:15)) {

TP [[hisp]] <- a %>% 
  filter(group_favorability_socialists != 999, pid3 <=3) %>%
  filter(hispanic==hisp) %>%
  group_by(age_group, pid3) %>%
  summarise(M = weighted.mean(group_favorability_socialists, na.rm=T)) %>%
  mutate(PID = recode(pid3, `1` = "Democrats", `2` = "Republicans", `3` = "Independents"),
          AGE = recode(age_group, `1` = "18-29 years old",
                            `2` = "30-44 years old",
                            `3` = "45-64 years old",
                            `4` = "65+ years old"
                            )) %>%
  mutate(row1 = row_number()) %>%
  unite(var, PID, AGE, sep = " who are ", remove = F)
}
```

In the general public, variation in ratings can be partly attributed to partisanship and age:

```{r}
# Make a  plot for everyone
a %>% 
  filter(group_favorability_socialists != 999, pid3 <=3) %>%
  group_by(age_group, pid3) %>%
  summarise(M = weighted.mean(group_favorability_socialists, na.rm=T)) %>%
  mutate(PID = recode(pid3, `1` = "Democrats", `2` = "Republicans", `3` = "Independents"),
          AGE = recode(age_group, `1` = "18-29 years old",
                            `2` = "30-44 years old",
                            `3` = "45-64 years old",
                            `4` = "65+ years old"
                            )) %>%
  mutate(row1 = row_number()) %>%
  unite(var, PID, AGE, sep = " who are ", remove = F) %>%
  ggplot(aes(x = M, y = reorder(var,-row1))) +
  geom_col(width = .25, fill = jzc[4]) +    coord_cartesian(xlim = c(1,4)) +
  theme_jz() +
  labs(title = "Ratings of socialists") +
  labs(x= "1 = Very favorable, 4 = Very unfavorable", y = "Group of respondents")
```

```{r, fig.height=7}
# list all categories
# a %>% count(hispanic)
  group <- NULL
  group[1] <- "Non-Hispanic"
   group[2] <-       "Mexican"
   group[3] <-         "Cuban"
   group[4] <-   "Argentinian"
   group[5] <-     "Colombian"
   group[6] <-    "Ecuadorian"
   group[7] <-   "Salvadorean"
   group[8] <-    "Guatemalan"
   group[9] <-    "Nicaraguan"
  group[10] <-     "Panamanian"
  group[11] <-       "Peruvian"
  group[12] <-        "Spanish"
  group[13] <-     "Venezuelan"
  group[14] <-  "Other Hispanic"
  group[15] <-  "Puerto Rican"


# To prepare a full set of charts, add print()
i <- 1 
for (hisp in c(1:5,12:15)) {
TP[[hisp]] %>%
  ggplot(aes(x = M, y = reorder(var,-row1))) +
  geom_col(width = .25, fill = jzc[i]) +    coord_cartesian(xlim = c(1,4)) +
  theme_jz() + 
  labs(title = "Ratings of socialists", subtitle = paste(group[hisp], "respondents")) +
  labs(x= "1 = Very favorable, 4 = Very unfavorable", y = "Group of respondents", caption = "Nationscape data")
  i<- i + 1
}

############################################################################
# Make a bar chart displaying avg. favorability by ideology, 
# age group, and a binary Hispanic indicator
############################################################################
howmany <- a %>% filter(group_favorability_socialists != 999, !is.na(ideo3)) %>%
                  filter(!is.na(age_group), !is.na(Hispanic)) %>%
                  tally() %>% pull()

a %>%
  filter(group_favorability_socialists != 999, !is.na(ideo3)) %>%
  group_by(age_group, ideo3, Hispanic) %>%
  summarise(M = weighted.mean(group_favorability_socialists, na.rm=T)) %>%
  mutate(HH = recode(Hispanic, `1` = "Hispanic", `0` = "non-Hispanic"),
          AGE = recode(age_group, `1` = "18-29 years old",
                            `2` = "30-44 years old",
                            `3` = "45-64 years old",
                            `4` = "65+ years old"
                            )) %>%
  unite(var, AGE, HH, ideo3, sep = " ", remove = F) %>%
  ggplot(aes(x = M, y = reorder(var,M))) +
  geom_point() +
  geom_segment(aes(x=0, xend=M, y=reorder(var,M), yend=reorder(var,M) ) ) +
  #geom_col(width = .44, fill = jzc[3]) +    coord_cartesian(xlim = c(1,4)) +
  theme_jz(fontsize = 12.5) + 
   coord_cartesian(xlim = c(1,4)) +
  labs(title = "Ratings of socialists") +
  labs(x= "1 = Very favorable, 4 = Very unfavorable", y = "Group of respondents", caption = paste("Nationscape data, N =",scales::comma(howmany)))
```

What about ratings of the police?

```{r, fig.height=6.5}
a %>%
  filter(group_favorability_the_police != 999, !is.na(ideo3)) %>%
  group_by(age_group, ideo3, Hispanic) %>%
  summarise(M = weighted.mean(group_favorability_the_police, na.rm=T)) %>%
  mutate(HH = recode(Hispanic, `1` = "Hispanic", `0` = "non-Hispanic"),
          AGE = recode(age_group, `1` = "18-29 years old",
                            `2` = "30-44 years old",
                            `3` = "45-64 years old",
                            `4` = "65+ years old"
                            )) %>%
  unite(var, AGE, HH, ideo3, sep = " ", remove = F) %>%
  ggplot(aes(x = M, y = reorder(var,M))) +
  geom_col(width = .44, fill = jzc[3]) +    coord_cartesian(xlim = c(1,4)) +
  theme_jz(fontsize = 12.5) + 
  labs(title = "Police favorability") +
  labs(x= "1 = Very favorable, 4 = Very unfavorable", y = "Group of respondents")
```

## Acknowledgments {.appendix}

Nationscape is designed and conducted by UCLA researchers led by political scientists Lynn Vavreck and Chris Tausanovitch Their transparency and fast data publication are gratefully acknowledged.