---
title: "Conservative and Liberal Labels: How much do they reveal about policy and candidate preferences?"
description: |
  Self-descriptions do not perfectly reflect citizens' policy positions (operational ideology) but conservative ideological labels correlate with support for conservative positions.
author:
  - name: Jan Zilinsky
date: 02-15-2021
output:
  distill::distill_article:
    self_contained: false
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
# Get data:
source("~/Dropbox/B_polls/Nationscape/R/0_load_and_recode_NS.R")

library(pollster)

## 'discrimination_blacks', 'discrimination_whites', 'discrimination_women',
##         'discrimination_men', 'discrimination_muslims', 'discrimination_christians']
```

Knowing people's self-ascribed ideological label seems to be quite useful for gauging where they stand operationally.

Specifically:

- Symbolic conservatives are more likely to oppose to Medicare for All than symbolic moderates and liberals.
- About 7 in 10 self-labeled conservatives disagree with abolishing private health insurance.
- People who call themselves conservatives are less likely to support raising the minimum wage.

```{r}
pollster::crosstab(df = a %>% filter(ideo5 <=5,
                                     medicare_for_all != 888),
                   x = ideo5,
                   y = medicare_for_all,
                   weight = weight, format = "long") -> medplot

obs <- nrow(a %>% filter(ideo5 <=5, medicare_for_all != 888))

medplot %>%
  ggplot(aes(x=ideo5,y=pct,fill=medicare_for_all)) + 
  geom_col(width = .4) + 
  coord_flip() +
  scale_fill_brewer(palette = 7, type = "qual") +
  labs(fill = "Enact Medicare-for-All?",
       x = "", 
       y = "Percent",
       title = "Distribution of support for Medicare for All by symbolic (self-reported) ideology",
       caption = paste("Data: Nationscape (2019-20).\nN = ", scales::comma(obs), ". Figure: @janzilinsky.", sep="")) +
  theme_bw()
  
```

```{r}
crosstab(df = a %>% filter(ideo5 <=5,
                                     abolish_priv_insurance != 888),
                   x = ideo5,
                   y = abolish_priv_insurance,
                   weight = weight, format = "long") -> abplot

obs2 <- nrow(a %>% filter(ideo5 <=5, abolish_priv_insurance != 888))

abplot %>%
  ggplot(aes(x=ideo5,y=pct,fill=abolish_priv_insurance)) + 
  geom_col(width = .4, alpha=.6) + 
  coord_flip() +
  scale_fill_brewer(palette = 2, type = "qual") +
  labs(fill = "Abolish private health insurance?",
       x = "", 
       y = "Percent",
       caption = paste("Data: Nationscape.\nN = ", scales::comma(obs2), ". Figure: @janzilinsky.", sep="")) +
  theme_bw()
```



```{r}
crosstab(df = a %>% filter(ideo5 <=5,
                                     minwage != 888),
                   x = ideo5,
                   y = minwage,
                   weight = weight, format = "long") -> mw

mw %>%
  ggplot(aes(x=ideo5,y=pct,fill=minwage)) + 
  geom_col(width = .4, alpha=.6) + 
  coord_flip() +
  scale_fill_brewer(palette = 2, type = "qual") +
  labs(fill = "",
       x = "", 
       y = "Percent",
       title="Raising the minimum wage to $15/hour?",
              caption = "Data: Nationscape. Figure: @janzilinsky.") +
  theme_bw()
```

```{r}
moe_crosstab_3way(df= a %>% filter(monthyear <= "2020-06-01",
                                   trump_biden != 999,
                                   minwage <=2),
                  x = trump_biden,
                  y = minwage,z=monthyear,
                  weight = weight,format="long",n=F,unwt_n=TRUE) %>%
  mutate(who = case_when(trump_biden=="Joe Biden" ~ "Biden supporters",
                         trump_biden=="Donald Trump" ~ "Trump supporters")) %>%
  filter(minwage == "Agree") %>%
  ggplot(aes(x=monthyear,y=pct,color=who)) + 
  geom_line(lwd=.5) + geom_point(size=1.5) +
  facet_grid(~who) +
  geom_errorbar(aes(ymin = (pct - moe), ymax = (pct + moe),
                    color=who), width=.1) +
  ylim(c(0,90)) +
  scale_color_brewer(palette = "Set1",direction = -1) +
  theme_bw() +
  labs(y="Percent",x="",
       title="Support for raising the minimum wage to $15/hour",
       subtitle="Democracy Fund + UCLA Nationscape data aggregated by month; undecided voters are excluded",
       caption = "Figure: @janzilinsky.") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=.5)) +
  theme(legend.position = "none") +
  scale_x_date(date_breaks = "1 month",date_labels = "%b 20%y")
```


```{r}
ECON <- tribble(
  ~econ_var,	~econ_name,
"Jgov_ins_agree",	"Provide government-run health insurance to all Americans",
"Jpublic_option_agree",	"Provide government health insurance option",
"Jhealth_subsidies_agree",	"Subsidize health insurance for low income Americans",
"Jraise_upper_tax_agree",	"Raise taxes on families making over $600,000",
"Jmctaxes_agree",	"Cut taxes for families making less than $100,000 per year",
"Jestate_NOT_repleal",	"Against eliminating the estate tax",
"Jcollege_agree",	"Provide debt-free state college",
"Jcap_agree",	"Cap carbon emissions to combat climate change",
"Jgreen_tech_agree",	"Invest in technology to protect the environment",
"Jguaranteedjob_agree",	"Guarantee jobs for all Americans",
"Jgreen_new_deal_agree",	"Enact a Green New Deal",
"Jmaternityleave_agree",	"Provide 12-week maternity leave",
"Jmedicare_agree",	"Enact Medicare-for-All",
"Jminwage_agree",	"Raise the minimum wage to $15/hour",
"Joil_drill_dereg_DISAGREE",	"Against removing oil and gas drilling barriers",
"Jreparations_agree",	"Grant reparations payments to descendants of slaves",
"Jright2W_DISAGREE",	"Against right-to-work laws",
"Jtrade_limit_agree",	"Limit trade with other countries",
"Juctaxes2_agree",	"Raise taxes on families making over $250,000",
"Jchina_tariffs_agree",	"Impose trade tariffs on Chinese goods",
"Jabolish_priv_insurance",	"Abolish private health insurance")

# a %>% select(all_of(econ_vars))

# Unweighted averages
# EE <- a %>% summarise_at(ECON$econ_var, mean, na.rm = TRUE) %>%
#   pivot_longer(everything(), names_to = "Issue") %>%
#   cbind(ECON$econ_name)

EE_all <- a %>% 
  summarise(across(ECON$econ_var, ~weighted.mean(., w = weight, na.rm = T))) %>%
  pivot_longer(everything(), names_to = "Issue") %>%
  cbind(ECON$econ_name)

EE_cons <- a %>% filter(ideo5 %in% c(4,5)) %>%
  summarise(across(ECON$econ_var, ~weighted.mean(., w = weight, na.rm = T))) %>%
  pivot_longer(everything(), names_to = "Issue") %>%
  cbind(ECON$econ_name)

# Numer of non-missing observations:
a %>% summarise(across(ECON$econ_var, ~sum(!is.na(.)))) %>%
  pivot_longer(everything(), names_to = "Non-missing") %>% pull(value) %>% mean()
```

```{r all-respondents}
EE_all %>% ggplot(aes(x=reorder(ECON$econ_name,value), y=value)) +
  geom_col(width = .9) +
  coord_flip() +
  labs(y="Percent holding a liberal/progressive position",
       x="", caption = "Data: Nationscape (2019-20).\nAverage N per item: 181,801") +
  scale_y_continuous(labels = scales::percent) +
  theme(text = element_text(size=14))
```

```{r conservative-respondents}
EE_cons %>% ggplot(aes(x=reorder(ECON$econ_name,value), y=value)) +
  geom_col() +
  coord_flip() +
  labs(y="Percent holding a liberal/progressive position",
       x="", subtitle = "Self-identified conservatives") +
  scale_y_continuous(labels = scales::percent)
```


```{r}
mean(EE_all$value)
mean(EE_cons$value)
```


```{r}
# moe_crosstab_3way(df= a %>% filter(registered==1,White==1,Hispanic==0,
#                                    monthyear <= "2020-06-01",
#                                    trump_biden != 999),
#                   x = votcat,
#                   y = trump_biden,z=monthyear,
#                   weight = weight,format="long",n=F,unwt_n=TRUE) %>%
#   ggplot(aes(x=monthyear,y=pct,color=trump_biden)) + 
#   geom_line(lwd=.5) + geom_point(size=1.5) +
#   geom_errorbar(aes(ymin = (pct - moe), ymax = (pct + moe),
#                     color=trump_biden), width=.1) +
#   labs(color="Preferred candidate",y="Two-party preference share",x="",
#        title="2020 vote intent among white non-Hispanic registered voters",
#        subtitle="Democracy Fund + UCLA Nationscape data aggregated by week; undecided voters are excluded") +
#   theme_minimal() +
#   facet_grid(~votcat) +
#   scale_color_brewer(palette = "Set1",direction = -1) +
#   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=.5)) +
#   scale_x_date(date_breaks = "1 month",date_labels = "%b 20%y")
```


